---
title: "Using Docker for deploying R APIs"
author: "David Mawdsley"
date: "3 March 2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

diamonds <- ggplot2::diamonds
```

## Introduction

* A walk through of using Docker to deploy an API written in R
* Example is a modified version of a phase 2 assist

* [Plumber](https://www.rplumber.io/) used to build API
* [Docker](https://www.docker.com/) used to deploy
* [nginx](https://www.nginx.com/) used to provide authentication and (eventually) https 

* (Bonus feature) - these slides are rendered using [Github Actions](https://github.com/features/actions) 


## Example data

Price of cut diamonds:

* Carat - weight of diamond
* Cut - quality of the cut (Fair, Good, Very Good, Premium, Ideal)
* Colour - colour of the diamond (D (best) to J (worst))
* Some other variables I didn't use

`r nrow(diamonds)` observations

(Based on model in [R for Data Science](https://r4ds.had.co.nz/model-building.html))

## Example data

```{r}
diamonds %>% 
  ggplot(aes(x = carat, y = price, colour = cut)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~color)
```

## Fitted model

```{r}

diamonds2 <- diamonds %>%
  mutate_if(is.factor, function(x) factor(x, ordered = FALSE)) %>%
  # Work with price on log scale
  mutate(lprice = log(price))


mod_diamond <- lm(lprice ~ carat + color + cut , data = diamonds2)


factor_levels <- diamonds2 %>%
  select_if(is.factor) %>%
  lapply(levels)
```

Linear model, using $\log(\mbox{price})$ as response:

```{r}
formula(mod_diamond)
```

$r^2=$ `r round(summary(mod_diamond)$r.squared,2)`


Haven't done any model checking - can undoubtedly do better.

## Model predictions

* Pass a data.frame containing new values to `predict()`

```{r, echo = TRUE}
predictvals <- tribble(~carat, ~color, ~cut,
                       1.5, "E", "Very Good") 
exp(predict(mod_diamond, predictvals))
exp(predict(mod_diamond, predictvals, interval = "pred"))

```

## Plumber


(Closest Python equivalents are probably [Flask](https://flask.palletsprojects.com/en/2.0.x/) or [FastAPI](https://fastapi.tiangolo.com/))

```{r, echo = TRUE, eval = FALSE}

#* Return the sum of two numbers
#* @param a The first number to add
#* @param b The second number to add
#* @post /sum
function(a, b) {
    as.numeric(a) + as.numeric(b)
}

```

* Generates Swagger docs automatically.
* R isn't ideal for APIs - single threaded.  
  * Various workarounds.


## Docker

* Container for code and runtime environment (_kind of_ like a VM)
* Self-contained

* Many pre-built images available; I used `rstudio/plumber`

## Dockerfile

```{r comment=''}
cat(readLines('../diamondpredict/Dockerfile'), sep = '\n')
```

## Docker - Build the container:

```{r engine='bash', comment='', eval = FALSE, echo = TRUE}
docker build -t diamondpredict .
```

Pulls in `rstudio/plumber`, and creates new "layers" as required. Cached.

And run:

```{r engine='bash', comment='', eval = FALSE, echo = TRUE}
docker run -p 127.0.0.1:8000:8000 diamondpredict 
```

`-p` option maps port 8000 in "dockerland" to port 8000 on the actual machine, listening on localhost





## Authentication

* Reverse proxy - forwards requests to API from outside world
* Using nginx to provide basic http authentication
  * Could use other authentication methods
* Can also be used to:
  * Set up https access
  * Load balance

`docker-compose` sets up nginx container and "dockerland" internal network


## Authentication (2)

```{r comment=''}
cat(readLines('../diamondpredict/docker-compose.yml'), sep = '\n')
```


## Moving to another system

Dump the container to a file:

```{bash, echo = TRUE, eval = FALSE}
docker save diamondpredict:latest |bzip2 > diamondpredict.tar.bz2
```

Move file to new system...

Load with:

```{bash, echo = TRUE, eval = FALSE}
docker load -i diamondpredict.tar.bz2
```

(Ideally you'd use a Docker Hub or a private Docker Registry)


## Summary

* Plumber provides an easy way for use to bolt an API onto our models
* Docker makes it easier to share and deploy the API
* Nginx lets us provide authentication, https, load balancing


## (Bonus) Github Actions

* R Markdown slides are rendered to html and published using Github Pages
* Github Actions:
  * Set up R environment (using `renv` $\approx$ Python's `venv`)
  * Render slides
  * Commit
  * Triggers Github pages "publish" workflow
  
